Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-19 providence/sections/intro.tex
Example ongoing

Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required. For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there might be no legal means of changing the outcome and this would significantly impact citizen confidence. For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome. This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size and hand count decisions.

@@

Definition \ref{def:RLA} is valid at the end of the RLA, and not at the end of each round. Thus it is not the case that an incorrect election will pass the audit if a sufficient number of rounds is drawn. In fact, the larger a sample the more accurate the estimate of election correctness can be, though, of course, it is possible to declare an incorrect outcome correct even after drawing a large sample.

We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}. Every time an audit stops, it incurs a contribution to its risk. Audit design determines how the risk contributions are distributed across rounds, so long as the overall risk is limited by $\alpha$.

@@

{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and they will keep drawing new rounds of ballots. See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins). Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.

EXAMPLE

@@

{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit. An audit is an RLA only if there is a proof of its risk-limiting property. Hence, at the very least, the adversary would need to invalidate an assumption of the proof to obtain:  
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0] > \alpha
$$

>>>>

Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required. For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there may not be a legal means of changing the outcome and this could significantly impact citizen confidence. For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome. This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size and hand count decisions.

@@

Definition \ref{def:RLA} is valid at the end of the RLA, and not at the end of each round. Thus it is not the case that an incorrect election will pass the audit if a sufficient number of rounds is drawn. In fact, the larger a sample, the more accurate the estimate of election correctness can be, and the more the audit will diverge from one that would stop (see the example below). Of course, though, it is possible to declare an incorrect outcome as correct even after drawing a large sample, but a good audit is less likely to make this error.

@@

\npara{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and new rounds will continue to be drawn. See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins). Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.

We present an example to illustrate the above idea. Consider the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with a total of 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.056614 in the pairwise contest (that is, ignoring the votes received by other candidates). The total number of votes cast in this contest, including invalid votes and votes for other candidates (all termed {\em irrelevant} votes for this pairwise contest), was 11, 315,056. A first round size of 2,217 corresponds to a stopping probability of 0.9 and a risk limit of 0.1 for \Providence (see \cite{usenix_minerva} for a description of how first round size may be computed, \Minerva and \Providence are identical in the first round). The stopping condition of \Providence (see section \ref{sec:prov}) is a comparison of the \Providence ratio with the risk limit, stopping when the ratio is smaller. We ran two simulations using these parameters.

Assuming the tally is correct, the probability that the audit stops in the first round is 0.9. Our simulation resulted in 1,138 votes for Trump, 1,054 for Biden and the rest were irrelevant votes. The \Providence ratio for this sample is 0.047, which is smaller than the desired risk limit, 0.1 and the audit stops.

Assuming the null, a tied election, the probability that the audit stops over all possible rounds is 0.1, the risk limit. Our simulation of a five-round audit did not stop. 
%The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09. 
Table \ref{tab:toy} lists the round sizes, votes for both candidates and the \Providence ratio after each draw.

Our simulation draws 1,111 and 1,079 votes for Trump and Biden respectively, to obtain a \Providence ratio of 0.256 which is not small enough. The simulation then proceeds to draw another round of cumulative size 5,970 for a conditional stopping probability of 0.9, which corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second). Similarly, this round corresponds to a risk expenditure of approximately 0.009. Trump gets 2,940 votes; Biden 2,953; and the ratio is 4.251. The process continues, with cumulative round sizes of 16,685; 35,096; and 76,979; Trump votes of 8,281; 17,320; and 37,943; Biden votes of 8,171; 17,264; and 37,868; and \Providence ratios of 3,150.268; 380,220,376; and $2.5e^{+22}$.

@@

\npara{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit. An audit is an RLA only if there is a proof of its risk-limiting property. Hence, at the very least, the adversary would need to invalidate an assumption of the proof to obtain:  
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0] > \alpha
$$
